Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex
more proof added to lemma 3.2

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $\deg(\lfloor{D}\rfloor)\ge 0$, so $R_D$ has a
generator $u$ in degree 1; let $u$ \todo{Was there supposed to be something else in this sentence?}.  Suppose $D$ is generated by $
u, x_1, \ldots, x_m$ in degree at most $\delta$ with relations generated 
in degree at most $\tau $.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $P \not \in \Supp(D), \alpha,\beta \in \mathbf{N}$ such
that

@@

Suppose $d\le \beta$ and $\frac{\bar{c}}{\bar{d}}\le \frac{\alpha}{
\beta}$ with $c>0$ is not a best lower approximation.  Then choose
best lower approximation $\frac{c_i}{d_i}$ such that $c_i\le c$ is
maximal; since $c<c_{i+1}$, $\frac{c_i}{d_i}\ge \frac{c}{d}$.  From
this we can deduce that $c_i+(d-d_i)\lfloor \frac{\alpha}{\beta}
\rfloor\ge c$.  Therefore, we can produce an element in $(R_{D'})_d$
 with pole degree $c$ at $P$ by either $y_i u^b$ or $y_j y_i u^b$
where $d_j=1$ and if $d_i=1$ as well then $c_i=1$. Thus, for a
general $d$ (not necessarily less than $\beta$), if $\frac{c}{d} \le
\frac{\alpha}{\beta}$ is not a best lower approximation than we can
construct an element in degree $d$ with pole degree $c$ by either $y
_n^a y_i u^b$ or $y_n^a y_j y_iu^b$ with the latter satisfying the
same conditions.  We notice that elements of this form are all
linearly independent in $R_{D'}$.  Thus, they form a $k$-basis for
$R_{D'}$ over $R_D$.  Thus $u,x_1,...,x_m,y_1,...,y_n$ generate
$R_{D'}$.

\todo{finish proof of relations, which follows easily using VZB's
method; peter will do this tomorrow (sunday)}

@@

\ssec{Saturation}
\label{ssec:g_0_saturation} \todo{Define saturation}
Here we compute saturation numbers. That is, given a list of
ramification orders, with $g = 0,\delta = 0$ understood, we have
the following saturations of the canonical divisor, as defined in 7.
2 of VZB \cite{vzb:stacky}. Exceptional cases are listed first
and generic cases follow.

@@

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq 2g - 1 + \max_{d \geq 0} \# S_{\sigma, J}(i, d)
		\end{align*}
		\todo{Check the 2g - 1, in particular, David seems to think we are usually 1 off and possibly 2 off when g = 1. I have a feeling this is a mistake, but don't understand why.} 
		where
		\begin{align*}
			S_{\sigma, J}(i, d) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i + 2d\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\label{thm:ram_order_ind}
Suppose $(\sx', 0, L')$ \todo{Peter: You should also instantiate $(\sx,0,L)$ as well} is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i})}{\deg (y_{i,e_i})} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\noindent
The assumption (Ad-iii) implies
\begin{align*}
	\deg \left( e_i L' - E_i \right) \geq 2g-1
\end{align*}

@@

By part (a), condition (Ad-ii), and Lemma
~\ref{lem:admissible_inequality} we have
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z)
	&\leq \frac{e_j-1}{2} + \deg z \left( \subhalf{e_j'} \right) -\frac{1}{e_j'} \\
	& = \frac{e_j-1}{2}-\frac{1}{e_j'}+\deg z \left( \subhalf{e_j'} \right) \\
	&= \frac{e_j(e_j-3)}{2(e_j-2)}+\deg z \left( \subhalf{e_j'} \right) \\
	&= \deg (y_{i,e_i}z) \subhalf{e_j'}.
\end{align*}

\vskip.1in
\noindent
{\sf Case 3: $j \in S(i,0).$}

In this case, we may first assume $z \neq y_{j,e_j}$, 
as this is covered by case 2, with $i$ and $j$ reversed. 
Hence, 
\begin{align*}
	-\ord_{Q_j}(z) \leq \deg z \subhalf{e_j'}
\end{align*}
implying
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z) 
	&\leq  e_i \subhalf{e_j'} + \deg z \subhalf{e_j'} 
	= \deg (y_{i,e_i}z) \subhalf{e_j'},
\end{align*}
completing part (c).

>>>>

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper})
satisfying $\deg(\lfloor{D}\rfloor)\ge 0$, so $R_D$ has a
generator $u$ in degree 1; let $u$.  Suppose $D$ is generated by $
u, x_1, \ldots, x_m$ in degree at most $\delta$ with relations generated 
in degree at most $\tau $.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $P \not \in \Supp(D), \alpha,\beta \in \mathbf{N}$ such
that

@@

Suppose $d\le \beta$ and $\frac{\bar{c}}{\bar{d}} \le \frac{\alpha}{
\beta}$ with $c>0$ is not a best lower approximation.  Then choose
best lower approximation $\frac{c_i}{d_i}$ such that $c_i\le c$ is
maximal; since $c<c_{i+1}$, $\frac{c_i}{d_i}\ge \frac{c}{d}$.  From
this we can deduce that $c_i+(d-d_i)\lfloor \frac{\alpha}{\beta}
\rfloor\ge c$.  Therefore, we can produce an element in $(R_{D'})_d$
 with pole degree $c$ at $P$ by either $y_i u^b$ or $z y_i u^b$ with $i>\lfloor \frac{\alpha}{\beta} \rfloor$, or $zu^b$
for some $z=y_j (y_{\lfloor \frac{\alpha}{\beta} \rfloor})^b$ where $d_j=1$ ($=d_{{\lfloor \frac{a}{b} \rfloor}}$) and $b<d_{\lfloor \frac{\alpha}{\beta} \rfloor + 1}$.  Thus, for a
general $d$ (not necessarily less than $\beta$), if $\frac{c}{d} \le
\frac{\alpha}{\beta}$ is not a best lower approximation than we can
construct an element in degree $d$ with pole degree $c$ by either $y
_n^a y_i u^b$, $y_n^a zy_iu^b$ or $y_n^a z u^b$ satisfying the
same conditions.  We notice that elements of this form are all
linearly independent in $R_{D'}$.  Thus, they form a $k$-basis for
$R_{D'}$ over $R_D$.  Thus $u, x_1, ..., x_m, y_1, ..., y_n$ generate
$R_{D'}$.

Choose $x_i\in \{x_1, \ldots, x_m\}$ and $y_j\in \{y_1, \ldots, y_n\}$, where $y_j$ corresponds to the best lower approximation $\frac{c_j}{d_j}$.  Recall from the previous paragraph that we can uniquely construct an element in degree $d$ with pole order $c$ at $P$ by elements of the $y_n^a y_l u^b$ and $y_n^a y_l y_{l'} u^b$ with $d_l=1$ and if $d_{l'}=1$ then $c_{l'} = \lfloor \frac{\alpha}{\beta} \rfloor$.  Specifically, we can successively subtract some $\gamma_l y_l u^{b_l}$, $\gamma_l y_l z_l u^{b_l}$, or $\gamma_l z_l$ with $y_l, z_l, b_l$ satisfying the same conditions as above and $\gamma_l\in k$ the element ensuring that the difference decreases in pole degree at $P$ by 1.  For convenience, we can combine these into the case $\gamma_l (y_l)^{s_l}(z_l)^{a_l}u^{b_l}$ with $(s_l,a_l)\in \{(1,0),(1,1),(0,1)\}$.  In the cases when $\frac{\alpha}{\beta}<1$ so $z_l$ is not well-defined, just pick an arbitrary element of $R_D$ for $z_l$ ensuring that the expression still makes sense; in such cases we will always choose $a_l=0$ so this choice is irrelevant, and only makes writing out cases more convenient.  Therefore, we can write
\[
	x_i y_j - \sum_{l} (y_l)^{s_l} (z_l)^{a_l} u^{b_l}\in R_D
\]
giving us a non-trivial relation with leading term $x_iy_j$. \todo{Explicitly state ordering at the beginning}

Similarly, suppose $y_i,y_j\in \{y_1, \ldots ,y_{n-1}\}$ with $1<d_i$ and $j\ge i$ (that is, so $y_iy_j$ does not appear as a basis element of the form described above: $y_n^0 y_jy_i u^0$).  Choose the maximal $l$ such that $c_h\le c_i+c_j$; since $1<d_i\le d_j$, $(c_i+c_j)-c_j>\lfloor \frac{\alpha}{\beta}\rfloor$ so $c_{j+1} < c_i + c_j$ meaning $c_h \ge c_{j+1} > c_j$.  Furthermore, $c_i + c_j - c_h<\lfloor \frac{\alpha}{\beta} \rfloor$ so either $c_h=c_i+c_j$ or there is some $h'$ such that $c_{h'}=c_i+c_j-c_l$ and $d_{h'}=1$.  Since $\frac{c_h}{d_h}$ is a best lower approximation of $\frac{\alpha}{\beta}$ with denominator larger than those of $\frac{c_i}{d_i}$ and $\frac{c_j}{d_j}$, we know that since $c_h\le c_i+c_j$ we must have $d_h<d_i+d_j$.  Thus in the case with a $y_{l'}$ term we have $d_h + d_{h'} = d_h + 1 \le d_i + d_j$.  That is to say, by choosing $b-d_i+d_j-ad_h-1$ we find (where $a\in\{0,1\}$ corresponding to if a $y_{h'}$ appears) that $y_h(y_h')^au^b$ and $y_iy_j$ both lie in degree $d_i+d_j$ with pole order $c_i+c_j$ at $P$.  From here we can follow a similar technique as in the previous paragraph to cancel out poles of $y_iy_j$ at $P$ to find
\[
	y_iy_j-\gamma_h (y_h)^{s_h}(z_{h})^{a_h}u^{b_h}-\sum_l (y_l)^{s_l} (z_{l})^{a_l}u^{b_l}\in R_D
\]
implying a relation with initial term $y_i y_j$.

Finally, suppose $y_i, y_j\in \{y_1, \ldots ,y_{n-1}\}$ such that $i\le j$, $d_j=d_i=1$, and $d_{j+1}=1$ (i.e. $c_i, c_j \le \lfloor \frac{\alpha}{\beta} - 1 \rfloor$).  If $i>1$ and then $y_{j+1}y_{i-1}$ lies in the same degree as $y_iy_j$ (degree 2) with the same pole order; following the same method as the previous two paragraphs we find
\[
	y_i y_j - \gamma y_{i-1} y_{j+1} - \sum_{l} (y_l)^{s_l} (z_l)^{a_l}u^{b_l}\in R_D.
\]
Similarly if $i=1$, then we get the same result using
\[
	y_i y_j - \gamma u y_{j+1} - \sum_l y_l u^{b_l} \in R_D.
\]
In both cases we deduce that $y_iy_j$ is an initial term of some relation.

Finally, we examine the case (for $\alpha \ge \beta$) of $(y_i)^{i + 1}$ when $i=\lfloor \frac{\alpha}{\beta}\rfloor$.  If $\beta=1$, then there are no relations with initial term a power of $y_i$ and we are immediately done.  Otherwise, assuming $\beta\ne 1$, we note that $c_id_{i+1}=c_{i+1}d_i+1=c_{i+1}+1$. \todo{reference to Evan's paper where he proves this}  Therefore, if $i>1$ then $c_i^{d_{i+1}+1}$ and $c_{i-1} c_{i+1}$ both lie in the $d_{i+1}+1$ degree with pole order $c_{i+1}+1$.  Following a similar technique to the previous paragraphs we find
\[
	(y_{i})^{c+1} - y_{i+1} y_{i-1} - \sum_l (y_l)^{s_l}(z_l)^{a_l} u^{b_l}
\]

Let $J$ be the ideal of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ generated by the initial terms described in the previous paragraphs.  We show that $J$ is in fact all of the initial ideal.  Suppose $f$ is a non-constant monomial that does not lie within $J$.  Further suppose $f$ does not lie in $k[u, x_1, \ldots, x_m]$.  Then there is some $y_i|f$.  Since $f\not\in j$, we cannot have some $x_j|f$ or else we would have $y_ix_j|f$ contradicting $f\not\in J$.  Furthermore, if $y_l|f$ for $l<i$, then we must have $d_l=1$.  If $d_i=1$ as well, then we must have $c_i=\lfloor \frac{\alpha}{\beta}\rfloor$, or else.  We observe, that such elements are precisely

@@

\ssec{Saturation}
\label{ssec:g_0_saturation}
Here we compute saturation numbers. That is, given a list of
ramification orders, with $g = 0,\delta = 0$ understood, we have
the following saturations of the canonical divisor, as defined in 7.
2 of VZB \cite{vzb:stacky}. Exceptional cases are listed first
and generic cases follow.

@@

\todo{Something is a little fishy with this order, it seems
something like: first grade by degree, then try to decrease the
highest y, if you cant, try to decrease the highest z, possibly
replace $z_5$ by $z_5+Cy_5$ to generic element?}

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq \# S_{\sigma, J}(i)
		\end{align*}
		where
		\begin{align*}
			S_{\sigma, J}(i) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i - e_j' \text{ for some } d \geq 0\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\label{thm:ram_order_ind}
Suppose $(\sx', 0, L')$ \todo{Peter: You should also instantiate $(\sx,0,L)$ as well} is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i'})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\noindent
The assumption (Ad-iii) implies
\begin{align*}
	\deg \left( e_i L' - E_i \right) \geq 0
\end{align*}

@@

By part (a), condition (Ad-ii), and Lemma
~\ref{lem:admissible_inequality} we have
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z)
	&\leq \frac{e_i-1}{2} + \deg z \left( \subhalf{e_i'} \right) -\frac{1}{e_i'} \\
	& = \frac{e_i-1}{2}-\frac{1}{e_i'}+\left( \subhalf{e_i'} \right) \\
	&= \frac{e_i(e_i-3)}{2(e_i-2)}+\left( \subhalf{e_i'} \right) \\
	&= \deg (y_{i,e_i}z) \subhalf{e_j'}.
\end{align*}

\vskip.1in
\noindent
{\sf Case 3: $j \in S(i,0).$}
In this case, we similarly obtain
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i})-\ord_{Q_j}(z) &\leq  e_i \subhalf{e_j'} +
	\deg z \subhalf{e_j'} = \deg (y_{i,e_i}z) \subhalf{e_j'}
\end{align*}
completing part (c).
